\section{Online MMKP}\label{sec_dynamic_seat}
Consider a set $\mathcal{N} = {1, 2, \ldots, N}$ of knapsacks, where each knapsack $j$ has a capacity $c_j \in \mathbb{Z}^{+}$. There is also a set $\mathcal{M} = {1, 2, \ldots, M}$ of distinct item types. Each item of type $i$ has a size $w_i \in \mathbb{Z}^{+}$ and yields a profit $r_i \in \mathbb{Z}^{+}$ when placed entirely into a knapsack. The item types are ordered such that their profit-to-weight ratios, $r_i / w_i$, are monotonically increasing in $i$.

Requests for these items arrive sequentially. Upon the arrival of a request (which specifies its type), the seller must immediately decide whether to accept or reject it. If accepted, the seller must also assign it to a specific knapsack with sufficient remaining capacity. Each item must be placed whole into a single knapsack; partial assignments or reassignments are not permitted.


To model this problem, we adopt a dynamic programming framework based on discrete time periods $t = 1, 2, \ldots, T$. In each period, at most one request arrives. Let $\lambda_i^t$ denote the probability that a request for an item of type $i \in \mathcal{M}$ arrives at time $t$. These probabilities satisfy $\sum_{i=1}^M \lambda_i^t \leq 1$ for all $t$, and we define $\lambda_0^t = 1 - \sum_{i=1}^M \lambda_i^t$ as the probability of no arrival in period $t$. Arrival events are assumed to be independent across time periods.


The system state is the vector of remaining capacities $\mathbf{C} = (c_1, \ldots, c_N)$. When an item of type $i$ arrives, the seller chooses a decision variable $u_{i,j}^t \in {0,1}$ for each knapsack $j$. The feasible set $U^t(\mathbf{C})$ is defined by:

$$
U^t(\mathbf{C})=\left\{u_{i, j}^t \in\{0,1\} \left\lvert\, \begin{array}{ll}
\text { (a) } & \sum_{j=1}^N u_{i, j}^t \leq 1 \quad \forall i \in \mathcal{M} \\
\text { (b) } & w_i u_{i, j}^t \leq c_j \quad \forall i \in \mathcal{M}, \forall j \in \mathcal{N}
\end{array}\right.\right\} .
$$

Constraint (a) ensures the item is assigned to at most one knapsack. Constraint (b) ensures that if the item is assigned to knapsack $j$ ($u_{i,j}^t=1$), its weight $w_i$ does not exceed $c_j$. The original vector form of this constraint, $w_{i}u_{i,j}^{t}\mathbf{e}_j \leq \mathbf{C}$ (where $\mathbf{e}_j$ is the $j$-th standard basis vector), is mathematically equivalent to (b).



Let $v^{t}(\mathbf{C})$ denote the value function, representing the maximum expected revenue obtainable from period $t$ onward, given the current remaining capacity vector $\mathbf{C}$. The Bellman equation is given by:

\begin{equation}\label{DP}
v^{t}(\mathbf{C}) = \max_{u_{i,j}^{t} \in U^{t}(\mathbf{C})}\left\{\sum_{i=1}^{M} \lambda_i^{t} \bigl( \sum_{j=1}^{N} r_i u_{i,j}^{t} + v^{t+1}(\mathbf{C} - w_i u_{i,j}^{t}\mathbf{e}_j)\bigr) + \lambda_0^{t} v^{t+1}(\mathbf{C})\right\}
\end{equation}

The boundary condition is $v^{T+1}(\mathbf{C}) = 0$ for all $\mathbf{C} \geq \mathbf{0}$, indicating that no more revenue can be earned after the final period $T$.

Let $\mathbf{C}_0 = (C_1, C_2, \ldots, C_N)$ be the initial capacity vector. The objective is to compute $v^{1}(\mathbf{C}_0)$, the maximum total expected revenue over the entire horizon from $t=1$ to $t=T$, and to find the policy of item assignments that achieves this value.


Solving the dynamic programming problem presented in Equation \eqref{DP} is computationally intractable for realistic problem sizes due to the curse of dimensionality inherent in the large state space.

To overcome this challenge, we propose a heuristic assignment policy. We first outline a traditional bid-price control policy. We then enhance this approach by introducing a novel bid-price control policy that leverages patterns.



% For any policy $\pi$, let $V^{\pi}$ denote the expected revenue collected under $\pi$. Among all policies, a special one is the dynamic programming (DP) policy as it achieves the maximal expected revenue. Apart from DP, another special policy is the hindsight optimum (HO), where the decision maker has full information on the demand realization for the entire time horizon and optimizes over the allocation schemes. Here, it is impossible for the HO to be attained because we can never know the full demand realization ahead. For ease of analysis, the hindsight optimum for the sample path is computed by solving a relaxed static problem and $V^{\text{HO}}$ is the expected value over all sample paths. Then for any online policy $\pi$, we have $$V^{\text{HO}} \geq V^{\text{DP}} \geq V^{\pi}.$$ It is well known that DP, despite its optimality, is computationally complex owing to the ``curse of dimensionality.'' Thus we use HO as our benchmark to evaluate the performance of any policy $\pi$.


\subsection{BPC Policy}
Bid-price control is a classical and widely studied methodology in network revenue management. The core idea is to set thresholds, known as bid prices, that represent the opportunity cost of consuming one unit of capacity. An item is accepted only if its revenue exceeds the estimated opportunity cost of the capacity it requires.

Typically, these bid prices are derived from the shadow prices of the capacity constraints in a deterministic approximation of the underlying stochastic problem. In this section, we detail the implementation of a bid-price control policy for our model.

We begin by formulating a deterministic linear programming (LP) approximation, specifically the LP relaxation of a multi-type multiple knapsack problem. This model uses expected demand over the horizon. Let $x_{ij}$ denote the number of type $i$ items assigned to knapsack $j$, and let $d_i = \sum_{t=1}^{T} \lambda^{t}_{i}$ represent the expected number of requests for type $i$. The formulation is as follows:


\begin{align}
\quad \max \quad & \sum_{i=1}^{M}  \sum_{j= 1}^{N} r_{i} x_{ij} \label{e0} \\
\text {s.t.} \quad & \sum_{j= 1}^{N} x_{ij} \leq d_{i}, \quad i \in \mathcal{M}, \label{deter_upper}\\ 
& \sum_{i=1}^{M} w_{i} x_{ij} \leq c_j, j \in \mathcal{N}, \label{capa_con} \\
& x_{ij} \geq 0, \quad i \in \mathcal{M}, j \in \mathcal{N}. \notag
\end{align}

The objective \eqref{e0} is to maximize total expected revenue. Constraint \eqref{deter_upper} ensures the total number of accepted type $i$ items does not exceed its expected demand. Constraint \eqref{capa_con} ensures the total weight in each knapsack $j$ does not exceed its initial capacity $C_j$.

The monotonic increase of the profit-to-weight ratio $r_i/w_i$ with type index $i$ implies that items with a higher index are more profitable per unit of capacity. Consequently, the optimal solution to the LP relaxation exhibits a greedy structure, preferentially utilizing higher-indexed item types. This structural property is formalized in Proposition \ref{sol_relax_deter}.

\begin{prop}\label{sol_relax_deter}
For the LP relaxation of the \textup{MMKP} problem, there exists an index $\tilde{i}$ such that the optimal solutions satisfy the following conditions: $x_{ij}^{*} = 0$ for all $j$, $i = 1,\ldots, \tilde{i}-1$; $\sum_{j=1}^{N} x_{ij}^{*} = d_{i}$ for $i = \tilde{i}+1,\ldots, M$; $\sum_{j=1}^{N} x_{ij}^{*} = \frac{L - \sum_{i = \tilde{i}+1}^{M} {d_i w_i}}{w_{\tilde{i}}}$ for $i = \tilde{i}$.
\end{prop}

The dual of LP relaxation of the MMKP problem is:

\begin{equation}\label{bid-price_dual}
    \begin{aligned}
    \min \quad & \sum_{i=1}^{M} d_i z_i + \sum_{j= 1}^{N} c_j \beta_{j} \\
    \text {s.t.} \quad & z_{i} + \beta_j w_i \geq r_i, \quad i \in \mathcal{M}, j \in \mathcal{N} \\
    & z_{i} \geq 0, i \in \mathcal{M}, \beta_{j} \geq 0, j \in \mathcal{N}.
    \end{aligned}
  \end{equation}

In \eqref{bid-price_dual}, $\beta_{j}$ can be interpreted as the bid-price for one size in knapsack $j$. A request is only accepted if the revenue it generates is no less than the sum of the bid prices of the sizes it uses. Thus, if $r_i -\beta_{j} w_i \geq 0$, meanwhile, the capacity allows, we will accept the item type $i$. And choose knapsack $j^{*} = \arg \max_{j} \{r_i -\beta_{j} w_i\}$ to allocate that item.

\begin{lem}\label{bid-price}
The optimal solution to problem \eqref{bid-price_dual} is given by $z_1 = \ldots = z_{\tilde{i}} =0$, $z_{i} = \frac{r_{i} w_{\tilde{i}} - r_{\tilde{i}} w_{i}}{w_{\tilde{i}}}$ for $i = \tilde{i}+1, \ldots, M$ and $\beta_j = \frac{r_{\tilde{i}}}{w_{\tilde{i}}}$ for all $j$.
\end{lem}

According to Lemma \ref{bid-price}, the decision inequality becomes $r_i -\beta_{j} w_i = r_{i} - \frac{r_{\tilde{i}}}{w_{\tilde{i}}} w_{i} \geq 0$. This establishes the threshold policy: reject item type $i, i < \tilde{i}$ and accept item type $i, i \geq \tilde{i}$.

\begin{algorithm}[H]
    \caption{Bid-Price Control}\label{algo_bid}
    \For{$t = 1, \ldots, T$}{
      {Observe a request of item type $i$\;}
      {Solve problem \eqref{bid-price_dual} with $\bm{d}^{[t,T]}$ and $\mathbf{C}^{t}$\;
      Obtain $\tilde{i}$ such that the aggregate optimal solution is $x e_{\tilde{i}} + \sum_{i=\tilde{i}+1} ^{M} d_{i}^{t} e_{i}$\;}
      \eIf{$i \geq \tilde{i}$ and $\max_{j \in \mathcal{N}}{c_j^{t}} \geq w_i$}
      {Set $k = \arg \min_{j \in \mathcal{N}}\{c_j^{t}|c_j^{t} \geq w_i\} $ and break ties\;
      Assign the item to knapsack $k$, let $c_{k}^{t+1} \gets c_{k}^{t} - w_{i}$ \;}
      {Reject the request\;}}
\end{algorithm}

% Let $val_{\theta}(I; \{d_{i}\})$ denote the optimal objective value of \eqref{theta_deter}.

% \begin{align}
%     \quad \max \quad & \sum_{i=1}^{M}  \sum_{j= 1}^{N} (n_i - \delta) x_{ij} \label{theta_deter} \\
%     \text {s.t.} \quad & \sum_{j= 1}^{N} x_{ij} \leq d_{i}, \quad i \in \mathcal{M}, \notag \\ 
%     & \sum_{i=1}^{M} n_{i} x_{ij} \leq \theta L_j, j \in \mathcal{N}, \notag 
% \end{align}

% Let $V_{\theta}^{OPT}(I)$ denote the expected value under optimal policy (relaxed) during $\theta T$ periods for instance $I$ (probability distribution).

% $V_{\theta}^{OPT}(I) = E_{\{d_{i}\}} [val_{\theta}(I; \{d_{i}\})] \leq val_{\theta} (I; \{E[d_{i}]\}) = val_{\theta}(I; \{\theta T p_{i}\})$

Let $z(\text{BPC})$, $z(\text{DLP})$ denote the optimal value of \eqref{bid-price_dual} and the LP relaxation of the \textup{MMKP} problem with expected demand, respectively. Then we have $z(\text{BPC}) = z(\text{DLP})$.

However, the BPC policy has two drawbacks. First, the capacity feasibility need to be checked when to accept a request. Second, when capacity permits, the policy treats all knapsacks as equally preferable, making no distinction among them.

\begin{example}
Consider $M =3$, $N =4$, $w_{1} = 3, w_{2} = 4, w_{3} = 5$, $r_{1} = 4, r_{2} = 6, r_{3} = 8$, $\bm{C} = [7, 8, 8, 4]$, $T = 8$, stationary arrival probability: $\lambda_{1} = \lambda_{3} = \frac{1}{4}$, $\lambda_{2} = \frac{1}{2}$. Then the expected demand for each type is $\bm{d} = (2, 4, 2)$.

For the traditional bid-price control, for each $j$, $\beta_{j}^{*} = \frac{4}{3}$. $z(BPC) = \frac{124}{3}$. Two drawbacks: there is no difference among knapsacks. Even $r_{3} - \beta_{4}^{*} * w_{3} > 0$, it is infeasible for type 3 assigned in knapsack 4.
\end{example}

\subsection{BPC Policy Based on Patterns (BPP)}
To account for the differences in placing items across knapsacks, we propose an enhanced dynamic programming (DP) formulation. The key idea, as detailed next, is to represent knapsack configurations using patterns rather than merely tracking the residual capacities.

A feasible pattern $\bm{h} = [h_{1}, \ldots, h_{M}]$ for knapsack $j$ satisfies $\sum_{i=1}^{M} w_{i} h_{i} \leq c_{j}$. Suppose that $S(c_{j})$ is the set of all feasible patterns for knapsack $j$. 

Let $v^t(\bm{C})$ denote the maximal expected value-to-go at time $t$, given the remaining capacity $\bm{C}$. The enhanced dynamic programming formulation is as follows:

\begin{equation}
v^t(\bm{C}) \geq \mathbb{E}_{i \sim \lambda^t}\left[\left\{
\begin{array}{ll}
\max \left\{\max\limits_{j:\bm{h} \in S(c_j), h_i \geq 1}\left\{v^{t+1}\left(\bm{C}-e_j^T \cdot w_i\right)+r_i\right\}, v^{t+1}(\bm{C})\right\},&\exists j, \text{satisfying} \bm{h} \in S(c_j), h_i \geq 1, \\
v^{t+1}(\bm{C}) & \text{otherwise}.
\end{array}\right]\right.
\end{equation}


The DP formulation involves two layers of maximization when there exists at least one knapsack $j$ satisfying $\bm{h} \in S(c_j), h_{i} \geq 1$. The inner maximization evaluates the optimal placement of item type $i$ across all feasible knapsacks $j$ where the pattern $\bm{h}$ is feasible for knapsack $j$ (i.e., $\bm{h} \in S(c_{j})$) and the item type $i$ can be accommodated (i.e., $h_{i} \geq 1$). The outer maximization compares the value of accepting $i$ (via the inner maximization) and rejecting $i$ (retaining $v^{t+1}(\bm{C})$). If no such $j$ exists, the request $i$ is rejected.


For notational convenience, we define $q_i$ as follows:
\begin{equation*}
    q_i= \begin{cases}\max _{j: \bm{h} \in S\left(c_j\right), h_i \geq 1}\left\{v^{t+1}\left(C-e_j^T w_i\right)+r_i\right\} & \text { if } \exists j \text{ satisfying } \bm{h} \in S\left(c_j\right), h_i \geq 1 \\ 0 & \text { otherwise }\end{cases}
\end{equation*}

We can solve the following program to compute $v^1(\bm{C})$ for any given capacity $\bm{C}$:

\begin{equation}\label{dp_bid}
    \begin{aligned}
    \min \quad & v^{1}(\bm{C}) \\
    \mathrm{s.t.} \quad & v^{t}(\bm{C}) \geq \mathbb{E}_{i \sim \lambda^t}\Bigg[\max\Big\{ q_{i}, v^{t+1}(\bm{C})\Big\}\Bigg], \\
    & v^{T+1}(\bm{C}) \geq 0.
    \end{aligned}
\end{equation}


Solving \eqref{dp_bid} remains computationally prohibitive. Following from the ADP approach, we approximate $v^{t}(\bm{C})$ as 

\begin{equation}\label{appro_dp}
    \hat{v}^{t}(\bm{C}) = \theta^{t} + \sum_{j=1}^{N} \max_{\bm{h} \in S(c_{j})} \{\sum_{i=1}^{M} \beta_{ij}^{\dag} h_{i}\}.
\end{equation}

The term $\beta_{ij}^{\dag}$ can be regarded as the approximated value for each type $i$ in knapsack $j$. Unlike traditional linear approximations, our approach retains the linear term $\theta^{t}$ but introduces a nonlinear component for each knapsack $j$. Specifically, we maximize the linear combination $\sum_{i=1}^{M} \beta_{ij}^{\dag} h_{i}$ over the feasible set $S(c_{j})$.

Our approximation extends classical linear ADP by incorporating resource-specific nonlinear terms through constrained maximization over feasible allocations. While similar separable corrections appear in resource allocation ADP (e.g., Powell, 2007), our explicit use of $\max_{\bm{h} \in S(c_j)}$ captures local constraints more directly.


% Our approximation adopts a separable nonlinear structure, but unlike classical approximations, the nonlinearity is implicitly defined by constrained maximization over feasible allocations. This captures problem-specific constraints.

Substituting \eqref{appro_dp} into \eqref{dp_bid}, we have:

\begin{align}
    \theta^{t} - \theta^{t+1} = \hat{v}^{t}(\bm{C}) - \hat{v}^{t+1}(\bm{C}) \geq \sum_{i} \lambda_{i}^{t} \max\left\{q_{i} - v^{t+1}(\bm{C}), 0\right\}
\end{align}

For the case where there exists a knapsack $j$ satisfying both conditions: $\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_j)} \sum_{i} \beta_{ij}^{\dag} h_{i}$ and $h_{i}^{*} \geq 1$, we establish the value difference: 

\begin{align*}
    & v^{t+1}(\bm{C} - e_{j}^{T} w_{i}) - v^{t+1}(\bm{C}) \\ 
  = & \max_{\bm{h} \in S(c_{j}- w_{i})} \{\sum_{i} \beta_{ij}^{\dag} h_{i}\} - \max_{\bm{h} \in S(c_{j})} \{\sum_{i} \beta_{ij}^{\dag} h_{i}\} \\
  = & -\beta_{ij}^{\dag} \leq 0
\end{align*}

The acceptance threshold is then defined as:
$$\alpha_{i} = \max\left\{\max_{j}\left\{r_i - \beta_{ij}^{\dag} \right\}, 0\right\},$$ 
with $\alpha_{i} = 0$ when no qualifying knapsack $j$ exists.

Let $\gamma_{j} = \max_{\bm{h} \in S(c_{j})} \{\sum_{i} \beta_{ij}^{\dag} h_{i}\}$. This yields:

\begin{align*}
    \theta^{1} & = \sum_{t=1}^{T} (\theta^{t} - \theta^{t+1}) \geq \sum_{t} \sum_{i} \alpha_{i} \lambda_{i}^{t} = \sum_{i} d_{i} \alpha_{i} \\
    \hat{v}^{1}(\bm{C}) & = \sum_{i} d_{i} \alpha_{i} + \sum_{j} \gamma_{j}
\end{align*}

Since $\hat{v}^{1}(\bm{C})$ constitutes a feasible solution to \eqref{dp_bid}, we have $\hat{v}^{1}(\bm{C}) \geq v^{1}(\bm{C}) =  V^{DP}$.

If all $j$ exist for $\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_j)} \sum_{i} \beta_{ij}^{\dag} h_{i}, h_{i}^{*} \geq 1$, the corresponding bid-price problem can be expressed as:

\begin{equation}\label{improve_bid}
    \begin{aligned}
    \min \quad & \sum_{i=1}^M \alpha_i d_i+ \sum_{j=1}^N \gamma_j \\
    \mathrm{s.t.} \quad & \alpha_i+ \beta_{ij}^{\dag} \geq r_i, \quad \forall i, j, \\
    & \sum_{i=1}^M \beta_{ij}^{\dag} h_i \leq \gamma_j, \quad \forall j, \bm{h} \in S(c_j), \\
    & \alpha_i \geq 0, \forall i, \quad \beta_{ij}^{\dag} \geq 0, \forall i, j \\
    & \gamma_j \geq 0, \quad \forall j.
    \end{aligned}
\end{equation}

$\alpha_{i}$ represents marginal revenue for type $i$. $\beta_{ij}^{\dag}$ represents the cost for type $i$ assigned in knapsack $j$. $\gamma_{j}$ represents the capacity cost associated with knapsack $j$.

When no knapsack $j$ exists satisfying both $\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_j)} \sum_{i} \beta_{ij}^{\dag} h_{i}$ and $h_{i}^{*} \geq 1$, the first set of constraints for $(i,j)$ should be removed from the formulation \eqref{improve_bid}. Although these constraints appear difficult to enforce in advance, the following lemma reveals that in practice we can avoid imposing additional restrictions. Simply solving problem \eqref{improve_bid} will inherently satisfy all required conditions.

% \begin{lem}
% For the optimal solution, $r_{i} \geq \beta_{ij}^{\dag}$.
% \end{lem}

\begin{lem}\label{BPP}
Define $\mathcal{J}_{i} = \{j \in \mathcal{N} \mid r_i - \beta_{ij}^{\dag *} \geq r_i - \beta_{ik}^{\dag *},\forall k \in \mathcal{N}, r_i - \beta_{ij}^{\dag *} >0 \}$.
If $\mathcal{J}_{i} = \varnothing$, the first set of constraints for $i$ is redundant.
If $\mathcal{J}_{i} \neq \varnothing$, then there exists a $j{'} \in \mathcal{J}_{i}$ such that:

$$\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_{j{'}})} \sum_{i} \beta_{ij{'}}^{\dag *} h_{i}, h_{i}^{*} \geq 1.$$
\end{lem}

This lemma guarantees that when such a $j{'}$ exists, the first set of constraints for $i$ becomes active; otherwise, it remains inactive. Consequently, we have $z(\text{BPP}) = \hat{v}^{1}(\bm{C})$.

Then, the control policy becomes as follow. If $\alpha_{i} > 0$, accept the type $i$. Find knapsack $k = \arg \max_{j \in \mathcal{N}}\{r_i - \beta_{ij}^{\dag}\}$, if multiple maximizers exist, assign the type $i$ to a knapsack $j$ satisfying: $$\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_j)} \sum_{i} \beta_{ij}^{\dag *} h_{i}, h_{i}^{*} \geq 1.$$ If $\alpha_{i} = 0$, check whether there exists a knapsack $j$ such that: $\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_j)} \sum_{i} \beta_{ij}^{\dag *} h_{i}, h_{i}^{*} \geq 1$. If no such $j$ exists, reject the type $i$; otherwise, accept the type $i$ and assign it to knapsack $j$.

Let $z(\text{BPC})$ and $z(\text{BPP})$ denote the expected optimal value of \eqref{bid-price_dual} and \eqref{improve_bid}, respectively.

\begin{lem}\label{BPC_relation}
    For the optimal $\beta_{j}^{*}$ in \eqref{bid-price_dual}, there exist optimal $\beta_{ij}^{\dag *}$ in \eqref{improve_bid} satisfying $\beta_{ij}^{\dag *} \leq w_{i} \beta_{j}^{*}$ for all $i$. Furthermore, $z(\text{BPC}) \geq z(\text{BPP})$.
\end{lem}
    
% $\beta_{ij}^{\dag} = w_{i} \beta_{j}$ is feasible for \eqref{improve_bid}, then the optimal bid-price 
    
Both bid-price approaches give upper bounds on the value function at any state, meanwhile it follows that BPP provides a tighter approximation to the value function more accurately than BPC does.

Under the approximation \eqref{appro_dp}, the BPP policy operates as follows:

% accept type $i$ if $r_i - \beta_{ij}^{\dag} > 0$ and at least one knapsack can assign type $i$, check  item type $i$ otherwise. 

% $\max_{j \in \mathcal{N}}{c_j^{t}} \geq w_i$

\begin{algorithm}[H]
    \caption{Bid-Price Control Based on Patterns}\label{algo_improve_bid}
    \For{$t =1, \ldots, T$}{
      {Observe a request of item type $i$\;}
      {Solve problem \eqref{improve_bid} with $\bm{d}^{[t, T]}$ and $\mathbf{L}^{t}$\;}
      \eIf{$r_i - \beta_{ij}^{\dag} > 0$}
      {Set $k = \arg \max_{j \in \mathcal{N}}\{r_i - \beta_{ij}^{\dag}\}$\;
      If multiple $k$s exist, assign the type $i$ to a knapsack $j$ satisfying: $$\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_j)} \sum_{i} \beta_{ij}^{\dag *} h_{i}, h_{i}^{*} \geq 1.$$
      Let $c_{j}^{t+1} \gets c_{j}^{t} - w_{i}$\;}
      {\eIf{There exists $j$ such that $\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_j)} \sum_{i} \beta_{ij}^{\dag} h_{i}, h_{i}^{*} \geq 1$}{Assign the item in knapsack $j$, let $c_{j}^{t+1} \gets c_{j}^{t} - w_{i}$\;}
      {Reject the request\;}}}
\end{algorithm}

% Furthermore, the BPP policy exhibits a threshold structure similar to the BPC policy. There exists a threshold type $\tilde{i}$ (or $\tilde{i}-1$, depending on the remaining capacity and expected future demand) such that all item types $i \leq \tilde{i}$ (or $i \leq \tilde{i}-1$) are rejected and larger types $i > \tilde{i}$ (or $i > \tilde{i}-1$) are accepted.

% Specifically, we have 
% \begin{lem}\label{BPP_decision}    
%     For the threshold index $\tilde{i}$ in Proposition \ref{sol_relax_deter},
%     \begin{itemize}
%         \item If $\sum_{i=\tilde{i}}^{M} d_{i} w_{i} > \sum_{j=1}^{N} c_{j}$, then $r_i - \beta_{ij}^{\dag} \leq 0$, for $i \leq \tilde{i}$,
%         \item If $\sum_{i=\tilde{i}}^{M} d_{i} w_{i} = \sum_{j=1}^{N} c_{j}$, then $r_i - \beta_{ij}^{\dag} \leq 0$, for $i \leq \tilde{i}-1$.
%     \end{itemize}
% \end{lem}

% Lemma \ref{BPP_decision} indicates how the relation between the remaining capacity and the expected demand affects the decision.

% when $\sum_{i=\tilde{i}}^{M} d_{i} w_{i} > \sum_{j=1}^{N} c_{j}$, item type $i, i \leq \tilde{i}$ is rejected and item type $i, i > \tilde{i}$ is accepted; when $\sum_{i=\tilde{i}}^{M} d_{i} w_{i} = \sum_{j=1}^{N} c_{j}$, item type $i, i \leq \tilde{i}-1$ is rejected and item type $i, i \geq \tilde{i}$ is accepted.

% Proof:
% $\sum_{j} x_{ij}^{*} < d_{i}$, then $\alpha_{i}^{*} = 0$. $i - \beta_{ij}^{\dag} \leq 0$.

% Lemma \ref{BPC_relation} shows that the I-BPC policy does not uniformly reject all small groups. Instead, it selectively accepts them based on pattern-based allocation, enabling more flexible decision-making. Unlike the traditional BPC, the I-BPC policy accounts for row-specific characteristics in allocation decisions. Therefore, the I-BPC policy better captures the structure of the capacity than the BPC policy does.


Unlike the traditional BPC, the BPP does not require explicit feasibility checks on capacity. However, it still needs to verify whether the optimal pattern contains the given request. This introduces a new challenge, as bid-price policies are inherently derived from a dual formulation, which may inherently omit key information preserved in the primal problem.

\begin{example}
Continue with the above example:

For the BPP, $\bm{\beta}_{1}^{*} = [4, 4, 4, 6]$, $\bm{\beta}_{2}^{*} = [6, 6, 6, 6]$, $\bm{\beta}_{3}^{*} = [10, 8, 8, 8]$. $z(BPP) = 40$. $\alpha = [0, 0, 0]$, $\gamma = [10, 12, 12, 6]$.

$r_{1} - \bm{\beta}_{1}^{*} = [0, 0, 0, -2]$, $r_{2} - \bm{\beta}_{2}^{*} = [0, 0, 0, 0]$, $r_{3} - \bm{\beta}_{3}^{*} = [-2, 0, 0, 0]$.

Drawback: need to check whether $j$ exists for $\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_{j_0})} \sum_{i} \beta_{ij_0}^{\dag *} h_{i}, h_{i}^{*} \geq 1$.

For example, $r_{3} - \bm{\beta}_{3}^{*} = [-2, 0, 0, 0]$ indicates type 3 cannot be assigned in knapsack 1. While the generated patterns for knapsack 4 are $[0, 1, 0], [1, 0, 0]$, which do not contain $h_{3}$, type 3 can only be assigned to knapsack 2 or 3.
\end{example}

While this verification is cumbersome under pure bid-price frameworks, the primal problem offers a more direct way to guarantee the existence of such a request-pattern match. To address this gap, we propose the dynamic primal formulation.

\subsection{Dynamic Primal Based on Patterns}

Let $y_{j \bm{h}}$ denote the proportion of pattern $\bm{h}$ used in knapsack $j$. The primal problem can be formulated as:

\begin{equation}\label{improve_primal}
    \begin{aligned}
    \max \quad & \sum_{i=1}^M \sum_{j=1}^N r_i x_{i j} \\
    \text {s.t.} \quad & \sum_{j=1}^N x_{i j} \leq d_i, \quad i \in \mathcal{M}, \\
    & x_{i j} \leq \sum_{\bm{h} \in S(c_{j})} h_i y_{j \bm{h}}, \quad i \in \mathcal{M}, j \in \mathcal{N}, \\
    & \sum_{\bm{h} \in S(c_{j})} y_{j \bm{h}} \leq 1, \quad j \in \mathcal{N}.
    \end{aligned}
\end{equation}

The first set of constraints demonstrate that for each item type $i$, the sum of assigned items and unassigned items equals the total demand. The second set of constraints shows that the number of items of type $i$ assigned in knapsack $j$ is not larger than the sum of $h_{i}$ (the count of type $i$ items in pattern $\bm{h}$) weighted by the pattern proportions $y_{i \bm{h}}$. The total proportion of patterns uesd in knapsack $j$ cannot exceed 1.

\begin{lem}\label{primal}
The optimal solution $x_{ij}^{*}$ to \eqref{improve_primal} satisfies $x_{ij}^{*} > 0$ if and only if there exists a knapsack $j$ such that $$\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_j)} \sum_{i} \beta_{ij}^{\dag} h_{i}, h_{i}^{*} \geq 1.$$
\end{lem}

In contrast to Lemma \ref{BPP}, this lemma demonstrates the equivalence between the condition $x_{ij}^{*} > 0$ and the existance of a knapsack $j$ satisfying $\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_j)} \sum_{i} \beta_{ij}^{\dag} h_{i}, h_{i}^{*} \geq 1$. This equivalence eliminate the need for explicit existance verification.

% Then, the following hierarchy holds: $V(\text{DLP}) = V(\text{BPC}) \geq V(\text{I-BPC}) \geq V^{\text{DP}}$.

% Any feasible solution to \eqref{improve_primal} yields a feasible solution to linear relaxation of SPDR having the same objective value, then we have  $V(\text{BPC}) \geq V^{\text{HO}} \geq V^{\text{DP}}$.

\begin{lem}
$z(\text{DLP}) \geq V^{\text{HO}}$ results from the concave property. 
\end{lem}

Consider the standard linear program: $\phi(\bm{d})= \{\max c^{T} \bm{x}: A \bm{x} \leq \bm{d}, \bm{x} \geq \bm{0}\}$.  Suppose that $\bm{d}_1$ and $\bm{d}_2$ are two demand vectors, the optimal solution is $\bm{x}_1$ and $\bm{x}_2$. For any $\lambda \in [0, 1]$, $\bm{d}_{\lambda} = \lambda \bm{d}_{1} + (1- \lambda) \bm{d}_{2}$. Let $\bm{x}_{\lambda} = \lambda \bm{x}_{1} + (1-\lambda) \bm{x}_{2}$, then $A \bm{x}_{\lambda} = A(\lambda \bm{x}_{1} + (1-\lambda) \bm{x}_{2}) \leq \lambda \bm{d}_{1} + (1- \lambda) \bm{d}_{2} = \bm{d}_{\lambda}$. Thus, $\bm{x}_{\lambda}$ is a feasible solution for $\bm{d}_{\lambda}$. Then, $\phi(\bm{d}_{\lambda}) \geq \bm{c}^{T} \bm{x}_{\lambda} = \lambda \bm{c}^{T} \bm{x}_{1} + (1-\lambda) \bm{c}^{T} \bm{x}_{2} = \lambda \phi(\bm{d}_{1}) + (1- \lambda) \phi(\bm{d}_{2})$, which indicates $\phi(\bm{d})$ is concave. Let $\phi(\bm{d})$ indicate the optimal value of the linear relaxation of the SPDR problem. Substitute $\bm{x}$ with $y_{j \bm{h}}$ and view $y_{j \bm{h}}$ as the decision variables, then the concave property still holds for $\eqref{improve_primal}$. $V^{\text{HO}} = E[\phi(\bm{d})] \leq \phi(E[\bm{d}]) = z(\text{DLP})$.

\subsubsection{Solve the dynamic primal}
% A pattern $\bm{h}$ is said to be inefficient if a mixture of other patterns can be used to generate more revenue for the same (or lower) consumption rate.

The pattern $\bm{h}$ is efficient for knapsack $j$ if and only if, for some $(\alpha_{1}, \ldots, \alpha_{M}, \gamma_{j})$ (except that $\alpha_{i} = r_i, \forall i$), $\bm{h}$ is the optimal solution to $$\max_{\bm{h}} \sum_{i=1}^{M} (r_i - \alpha_{i}) h_{i} - \gamma_{j}$$

To generate all efficient patterns, we need to solve the subproblem for each knapsack $j$:

\begin{align}\label{subproblem}
    \max \quad & \sum_{i=1}^{M} (r_i - \alpha_{i}) h_{i} - \gamma_{j} \\
    \text {s.t.} \quad & \sum_{i= 1}^{M} w_{i} h_{i} \leq c_j, \notag \\
    & h_{i} \in \mathbb{N}, \quad i \in \mathcal{M}. \notag
\end{align} 

If the optimal value of \eqref{subproblem} is larger than $0$, the primal $\eqref{improve_primal}$ reaches the optimal. Otherwise, a new pattern can be generated.

One important fact is that only efficient sets are used in the solution to \eqref{improve_primal}. Specifically, 

\begin{lem}\label{efficient}
    If $y_{j \bm{h}}^{*} > 0$ is the optimal solution to \eqref{improve_primal}, then $\bm{h}$ is an efficient pattern.
\end{lem}


A pattern $\bm{h}$ is dominant if there is no distinct pattern $\bm{h}{'}$ where every component of $\bm{h}{'}$ is greater than or equal to the corresponding component of $\bm{h}$. The efficient pattern is a dominating pattern. (If $\alpha_{i} = r_i$, \eqref{improve_primal} reaches the optimal and no pattern will be generated.)

The relation between the capacity and the demand shows the different structure of the optimal solution.

\begin{lem}
When $\sum_{i=1}^{M} d_{i} w_{i} < \sum_{j=1}^{N} c_{j}$, we have $\gamma_{j}^{*} =0, \forall j$, $\beta_{ij}^{\dag *} =0, \forall i,j$ and $\alpha^{*}_{i} = r_i, \forall i$. There exists at least one knapsack $j$ such that $\sum_{\bm{h} \in S(c_{j})} y_{j \bm{h}}^{*} < 1$.

When $\sum_{i=1}^{M} d_{i} w_{i} \geq \sum_{j=1}^{N} c_{j}$, we have $\sum_{\bm{h} \in S(c_{j})} y_{j \bm{h}}^{*} = 1, \forall j$.
\end{lem}


\begin{algorithm}[H]
    \caption{Dynamic Primal}\label{algo_improve_primal}
    \For{$t = 1, \ldots, T$}{
      {Observe a request of type $i$\;}
      \If{$c_{j}^{t} = w_i, \exists j$}
      {Assign the item to knapsack $j$\; 
      \Continue}
      {Solve problem \eqref{improve_primal} with $\bm{d}^{[t, T]}$ \;
      Obtain an optimal solution $x_{ij}$ \;}
      \eIf{$\max_{j}\{x_{ij}\} > 0$}
      {Set $k = \arg \max_{j}\{x_{ij}\}$ and break ties\;
      Assign the item to knapsack $k$, let $c_{k}^{t+1} \gets c_{k}^{t} - w_{i}$\;}
      {Reject the request\;}}
\end{algorithm}

Meanwhile, it guarantees feasible placement. Once a request is accepted, the policy ensures it can be assigned to a suitable knapsack without additional feasibility checks.

\begin{example}
For the primal, $\bm{x}_{1}^{*} = [1, 1, 0, 0]$, $\bm{x}_{2}^{*} = [1, 0, 2, 1]$, $\bm{x}_{3}^{*} = [0, 1, 0, 0]$. It indicates that type 1 can be assigned in knapsacks 1 or 2, type 2 cannot be assigned in knapsack 2, type 3 can only be assigned in knapsack 2. 

It may contain multiple efficient patterns for one knapsack. In this example, there is exactly one efficient pattern for each knapsack: $[1, 1, 0]$, $[1, 0, 1]$, $[0, 2, 0]$, $[0, 1, 0]$. It shows that knapsack 1 can assign type 1 and 2, so on and so forth.

Using $x_{ij}$ to make the decision is straightforward and easy to implement.
\end{example}


\begin{example}
Consider $M =3$, $N =4$, $w_{1} = 3, w_{2} = 4, w_{3} = 5$, $r_{1} = 4, r_{2} = 6, r_{3} = 8$, $\bm{C} = [7, 8, 8, 4]$, $T = 8$, stationary arrival probability, $\lambda_{1} = \lambda_{3} = \frac{1}{4}$, $\lambda_{2} = \frac{1}{2}$. Then the expected demand for each type is $\bm{d} = (2, 4, 2)$.

For the traditional bid-price control, for each $j$, $\beta_{j}^{*} = \frac{4}{3}$. $z(BPC) = \frac{124}{3}$. Two drawbacks: there is no difference among knapsacks. Even $r_{3} - \beta_{4}^{*} * w_{3} > 0$, it is infeasible for type 3 assigned in knapsack 4.

For the BPP, $\bm{\beta}_{1}^{*} = [4, 4, 4, 6]$, $\bm{\beta}_{2}^{*} = [6, 6, 6, 6]$, $\bm{\beta}_{3}^{*} = [10, 8, 8, 8]$. $z(BPP) = 40$. $\alpha = [0, 0, 0]$, $\gamma = [10, 12, 12, 6]$.

$r_{1} - \bm{\beta}_{1}^{*} = [0, 0, 0, -2]$, $r_{2} - \bm{\beta}_{2}^{*} = [0, 0, 0, 0]$, $r_{3} - \bm{\beta}_{3}^{*} = [-2, 0, 0, 0]$.

Drawback: need to check whether $j$ exists for $\bm{h}^{*} \in \arg\max_{\bm{h} \in S(c_{j_0})} \sum_{i} \beta_{ij_0}^{\dag *} h_{i}, h_{i}^{*} \geq 1$.

For example, $r_{3} - \bm{\beta}_{3}^{*} = [-2, 0, 0, 0]$ indicates type 3 cannot be assigned in knapsack 1. While the generated patterns for knapsack 4 are $[0, 1, 0], [1, 0, 0]$, which do not contain $h_{3}$, type 3 can only be assigned to knapsack 2 or 3.

% $[1, 1, 0], [2, 0, 0], [0, 0, 1]$, $[0, 2, 0], [1, 0, 1]$, $[0, 2, 0], [1, 0, 1]$, $[0, 1, 0], [1, 0, 0]$

For the primal, $\bm{x}_{1}^{*} = [1, 1, 0, 0]$, $\bm{x}_{2}^{*} = [1, 0, 2, 1]$, $\bm{x}_{3}^{*} = [0, 1, 0, 0]$. It indicates that type 1 can be assigned in knapsacks 1 or 2, type 2 cannot be assigned in knapsack 2, type 3 can only be assigned in knapsack 2. 

It may contain multiple efficient patterns for one knapsack. In this example, there is exactly one efficient pattern for each knapsack: $[1, 1, 0]$, $[1, 0, 1]$, $[0, 2, 0]$, $[0, 1, 0]$. It shows that knapsack 1 can assign type 1 and 2, so on and so forth.

Using $x_{ij}$ to make the decision is straightforward and easy to implement.
\end{example}

% Multiple subset sum

\subsection{Type Analyses}

When there is only one type, the optimal policy is straightforward, i.e., accept the request until the capacity is insufficient. 

When there are two types, 

When will the threshold policy be the optimal?