% !TEX root = sum1.tex
\section{Dynamic Demand}\label{dynamic_demand}

We also study the dynamic seating plan problem, which is more suitable for commercial use. In this situation, customers come dynamically, and the seating plan needs to be made without knowing the number and composition of future customers. 

It becomes a sequential stochastic optimization problem where conventional methods fall into the curse of dimensionality due to many seating plan combinations. To avoid this complexity, we develop an approach that aims directly at the final seating plans. As stated above, we can obtain such a seating plan from stochastic programming. However, it only gives the initial seat assignment without handling the dynamic situation. To further know how to make decisions when the group arrives dynamically, we discuss two different dynamic situations in the following section. 

% Specifically, we define the concept of target seating plans deemed satisfactory. In making the dynamic seating plan, we will try to maintain the possibility of achieving one of the target seating plans as much as possible.

\subsection{Dynamic Demand Situations}
Generally speaking, there are two ways to purchase tickets for concerts or movies: no seat selection when booking and seat selection when booking. We consider the following dynamic demand situations, ticket reservation without seat selection and seat assignment before arrival. % ticket reservation with row selection

Since singing concert venues have more seats and higher demand for tickets, organizers usually do not take the approach of determining the seats at the time of booking, but only provide the seat sections that can be selected at the time of booking or simply do not provide seat information, and then inform customers of the seat information after the overall seat layout is determined. Movie theaters or small concert, however, have relatively few seats, the attendance rate is usually low enough to allow free selection of seats directly online.


For reservations without seat selection, the seat assignment will not be made immediately, decision-maker only needs to reject or accept each request during the making-reservation stage. After the reservation deadline, the seat layout information will be informed to the customers before admission. 
(Row/second selection for extention)
% For example, some theaters only provide seat reservations for the audience.

For seat assignment before arrival, the specific procedure will be changed under the requirement of social distancing. The seat assignment will be arranged before the groups book the tickets, the groups only need to find the corresponding-size seats when accepting services. For example, prior seat planning can satisfy the requirement of social distancing and also save costs without changing seat allocation. The seat assignment could remain for one day because the same film genre will attract the same feature of different group types.
(Online for extention)
% , the movie halls can seal the empty seats to prevent people from taking seats.

% For reservation with row selection, 
% the concerts will allow people to choose the section and row when making the reservation.
% 根據客戶交易的日期及時間以先後次序分配，不設自行選擇座位
% 预订票指的是，第一类，预订门票并付完款之后，无需选座，大麦网会按照付款先后顺序配送门票；第二类，预订门票并付完款之后，需要选座的场次，大麦也会提前短信通知您所在分组的选座时间，准时参加即可

% https://zhuanlan.zhihu.com/p/39125507
% 优先购票  公开发售
% 由于位置必须相连，相连座位门票一般会比单张门票位置稍差，如果是单人购票，建议把相连座位的勾去掉，这样系统会按从近到远的顺序自动为你筛选出较近的门票。

%  为了公平 分阶段 

% We don't care about the arrival sequence; only the number of groups matters. Because as long as the approximation about the number of groups is accurate, we can handle any sequence.

% \subsection{Generate Scenarios with Stochastic Information}

\subsection{Mapping sequences to scenarios by discrete periods}\label{MappingSeq}
We can use stochastic information to generate scenarios integrated from sequences. Suppose there are $T$ independent periods; at most, one group will arrive in each period.

There are $J$ different group types(including the group with no people). Let $\mathbf{y}$ be a discrete random variable indicating the number of people in the group. Let $\mathbf{p}$ be the vector probability, where $p(y = j) = p_j$, $j = 0,1,\ldots,J-1$ and $\sum_{j} p_{j} =1$. Then a sequence can be expressed as $\{y_{1}, y_{2}, \ldots, y_{T}\}$. (It can be modeled as a multinomial distribution, $p(\mathbf{Y} \mid \mathbf{p})=\prod_{j=0}^{J-1} p_j N_j$).

Let $N_{j} = \sum_{t} I(y_t = j)$, i.e., the count number of times group type $j$ arrives during $T$ periods. Then the set of counts $N_{j}$ (scenarios) follows a multinomial distribution, $$p\left(N_0, \ldots, N_{J-1} \mid \mathbf{p}\right)=\frac{T !}{N_{0}!, \ldots, N_{J-1}!} \prod_{j=0}^{J-1} p_{j}^{N_j}, T = \sum_{j=0}^{J-1} N_{j}$$

% scenario:

% sequence will affect the result.
% Then, if we fix the seat assignment as $[3,3]$. For the sequence, $2,2,2,2,2,3,3,3,3,3$, we will place one $2$ in a three-seat. 
% But for the sequence, $3,3,3,3,3,2,2,2,2,2$, we will accept three $3$ and three $2$.

Show the complexity: the number of different sequences $J^{T}$, and the number of scenarios is $O(T^{J-1})$.

Obtained by DP:
Use $D(T,J) $ to denote the number of scenarios, which equals to the number of different solutions to $x_{1}+\ldots + x_{J} = T, \mathbf{x} \geq 0$.
Then we know the recursion relation $D(T, J) = \sum_{i= 0}^{T} D(i, J-1)$ and $D(i,2) = i+1, D(i,1) = 1$.
$D(T,3) = \frac{(T+2)(T+1)}{2}, D(T,J) = O(T^{J-1})$.

The number of scenarios is too large to enumerate all possible cases.
Thus, we choose to sample some sequences from the multinomial distribution.

Remark: this approach still works under the assumption that time is continuous.
% Example:
% The group types are $[2,3,4,5]$. The number of periods is $20$. The number of given rows is 4 and the number of seats is 22.
% Each group arrives with the same probability.
% The number of sequences generating from multinomial distribution is $1000$. Then, we can obtain $[0,3,6,11]$ from stochastic programming. When the number of sequences is 5000, we still obtain $[0,3,6,11]$. It shows that sampling is practical.

\subsection{Nested Policy with Given Supply}\label{nested_policy}
% Recall that the stochastic programming only considers the situation that small-size groups can use the surplus large-size seats.

Once we obtain a solution from the stochastic programming, we need to follow some basic rules to assign the seats.
\begin{itemize}
    \item When the supply of one arriving group is enough, we will accept the group directly.
    \item When the supply of one arriving group is 0, the demand can be satisfied by only one larger-size supply.
    \item When one group is accpected to occupy the larger-size seats, the rest empty seat(s) can be reserved for the future demand.
\end{itemize}

We can assign the seats to the corresponding-size group. But when a group comes while the corresponding supply is 0, should we give this group to the larger-size seats? Now we demonstrate the nested policy for this problem.

Suppose we accept a group of $i$ to take over $j$-size seats. In that case, the expected served people is $i + (j-i-1)P(D_{j-i-1} \geq x_{j-i-1}+1)$, where $i < j$, $P(D_i \geq x_i)$ is the probability of that the expected demand of group type $i$ in the following periods is no less than $x_i$, the remaining supply of group type $i$.

When a group of $i$ occupies $j$-size seats, $(j-i-1)$ seats can be provided for one group of $j-i-1$ with one seat of social distancing.
Thus, the term, $P(D_{j-i-1} \geq x_{j-i-1}+1)$, indicates the probability that the demand of group type $(j-i-1)$ in the future is no less than its current remaining supply plus 1. If $j -i-1 =0$, then this term equals 0.

Similarly, when the expected demand of a group of $j$ in the future is no less than its remaining supply currently, we would reject a group of $i$, the expected served people is $j P(D_{j} \geq x_{j})$.

Let $d(i,j)$ be the difference of expected served people between acceptance and rejection on group $i$ occupying $j$-size seats. Then $d(i,j) = i + (j-i-1)P(D_{j-i-1} \geq x_{j-i-1}+1) - j P(D_{j} \geq x_{j}), j >i$.

One intuitive decision is to choose the largest difference.
We can obtain $d(i,j) = j P(D_{j} \leq x_{j} -1) - (j-i-1)P(D_{j-i-1} \leq x_{j-i-1}) -1$ after reformulating. 
Let $F_{j}(x;T)$ be the cumulative distribution function of the number of arrival groups $D_{j}$ in $T$ periods. Then $F_{j}(x; T_{r}) = P(D_{j} \leq x)$, and $D_{j}$ follows a binomial distribution $B(T_{r}, p_{j})$, where $T_{r}$ is the numebr of remaining periods.

Thus, $d(i,j) = j F_{j}(x_{j}-1; T) - (j-i-1) F_{j-i-1}(x_{j-i-1}; T) -1$. For all $j >i$, find the largest $d(i,j)$, denoted as $d(i,j^{*})$.

If $d(i,j^{*}) >0$, we will place the group $i$ in $j^{*}$-size seats. Otherwise, reject the group.

The algorithm is shown below:

\begin{algorithm}[H]
  \caption{Nested policy under given supply}\label{algo_nested_policy}
  \begin{description}
    \item[Step 1.] Obtain a supply, $\X^{0} = [x_1,\ldots,x_{J}]$, from the stochastic programming.
    \item[Step 2.] For the arrival group type $i$ at period $T{'}$, if $x_{i} > 0$, accept it. Let $x_{i} = x_{i} -1$. Go to step 4.
    \item[Step 3.] If $x_{i} = 0$, find $d(i,j^{*})$. If $d(i,j^{*})>0$, accpect group type $i$. Set $x_{j^{*}} = x_{j^{*}} -1$. Let $x_{j-i-1} = x_{j-i-1} + 1$ when $j-i-1>0$. If $d(i,j^{*}) \leq 0$, reject group type $i$.
    \item[Step 4.] If $T{'} \leq T$, move to next period, set $T{'} = T{'}+1$, go to step 2. Otherwise, terminate this algorithm.
  \end{description}
\end{algorithm}

% Example:
% If the supply is $[0,3,6,11]$, then here comes a group of 1. There will be three choices.
% \begin{align*}
% 1 \geq 2 P(D_{2} & \geq x_{2}) \\
% 1 + 1\cdot P(D_{1}\geq 1) & \geq 3 P(D_{3}\geq x_{3}) \\
% 1 + 2\cdot P(D_{2}\geq (1+ x_{2})) & \geq 4 P(D_{4}\geq x_{4})
% \end{align*}
% $\mathbf{x}$ is the remaining supply right now.

We show the results of Benders and IP under nested policy in section \ref{Bender_IP}.


\subsection{Dynamic Situation with Stochastic Information}
In this section, we will present methods to assign seats with stochastic information.


\subsubsection{Ticket Reservation without Seat Selection}
We develop a DP-based method by relaxing all rows to one row with the same capacity.
Suppose there always exists one seat assignment under the total capacity. Then we can use DP to make the decision in each period.

$$V(S,T) = \sum_{i \in N} p_i \max\{ {[V((S-s_i-1),T-1)+ s_i]}, {V(S,T-1)}\}$$

After obtaining the pre-accepted sequence, we still need to check whether this sequence is feasible for the seat assignment. In most cases, it is feasible. That is the reason why we use relaxation. When it is not feasible, we should delete the group one by one from the last arrival of the sequence until it is feasible. In practice, we reject the request in the pre-accepted sequence until we cannot find a feasible seat assignment.

% We need to assign seats to the group for each arrival. In each period, the group can select the row they want to sit when the capacity is enough. FCFS will be more appropriate. But M1 and M3 can also be used.

\subsubsection{Seat Assignment before Arrivals}
% we make the decision according to the realized arrivals.
The stochastic seat assignment(SSA) method can give a seat assignment before the group arrivals. The specific procedures are demonstrated in the above sections. The brief steps are shown below: 

\begin{algorithm}[H]\label{general_method}
  \caption{Stochastic seat assignment method}
  \begin{description}
    \item[Step 1.] Obtain a linear solution from stochastic programming \eqref{BD_master1}.
    \item[Step 2.] Obtain the near-optimal integral solution from deterministic model with lower and upper bounds.
    \item[Step 3.] Accept or reject group arrivals according to the nested policy in section \ref{nested_policy}.

    % update the scenario and the probability, add constraints when re-calculating stochastic programming.
    % we can update the supply whenever some demand exceeds the supply.
  \end{description}
\end{algorithm}

% Then use Algorithm \ref{algo_nested_policy} to make the decision.


% \subsubsection{Ticket Reservation with Row Selection}
% There are two methods to achieve this goal.

% First, we can generate patterns planning according to stochastic information. Calculate the maximal supply from the mean demand. The supply gives the number of patterns with different losses; then we prepare the corresponding pattern planning for each row. Every group will be assigned in each period according to the designated row as long as the capacity allows. 

% The number of combinations is enormous.

% The second one is named the seat row selection method based on the stochastic seat assignment method. The initial supply is obtained from the stochastic model, then update the supply from the deterministic model after the first period. After accepting one group, we update the accepted demand and remaining seats. The policy follows section \ref{nested_policy}.


% Set the mean demand as the initial supply, update the supply from deterministic model by setting accpected demand as the lower bound. / can be used in scenario 1,2.



% Partially dynamic: at the beginning stage, the capacity is sufficient, thus we will accept all arrivals. 
% Multiple planning approach 

\subsection{Benchmark}
 
The benchmarks for the above two situations are described below.

\subsubsection{FCFS}\label{FCFS-based}
For ticket reservation without seat selection, the intuitive but trivial method will be on a first-come-first-served basis. Relax all rows to one row with the total number of seats. For the arrival sequence, find the target arrival when the number of seats taken by the preceding arrivals does not exceed the capacity. 
% And use the nested policy to accept or reject one group in the remaining arrivals. 
Then we obtain a new sub-sequence, including the arrivals from the first to the target and a possible arrival. For the convenience of calculation, we check the feasibility of constructing a seat assignment from the end of the sub-sequence. When it is not feasible for the seat assignment, we should delete the group one by one from this sub-sequence until a feasible seat assignment is found. In reality, we need to check the feasibility one group by one. 

% Each request will be assigned row by row. When the capacity of one row is not enough for the request, we arrange it in the next row. If the following request can take up the remaining capacity of some row exactly, we place it in that row immediately. We check each request until the capacity is used up. 


\subsubsection{Largest Patterns Planning}\label{largest_pattern}
For each row, we choose the patterns from $I_1$. Accept the group such that the largest pattern can be maintained. When the arrival cannot be assigned in the planning patterns from $I_1$, we can change the largest pattern to a second largest pattern according to the coming arrival.


% There is a reservation stage, we only decide to accept or reject.
% After certain periods, there is a seat selection stage.

% Or use partial static information to estimate the probabilities, then generate new plannings.

% Multiple scenario approach:
% Suppose that we know the probabilities, we can use the sampling demands to estimate which patterns we should use. For example, three rows with $I_1$, seven rows with $I_2$. For each arrival, if there exists one scenario containing this group, we accept it; otherwise we use nested policy to accept or reject it.

\begin{algorithm}[H]\label{algo_largest}
  \caption{Method by using the largest patterns}
  \begin{description}
    \item[Step 1.] Generate the largest pattern by the greedy way for each row.
    \item[Step 2.] Denote the minimal and maximal size of group in the pattern of row $i$ as $\min_i$ and $\max_i$. 
    \item[Step 3.] For the arrival with the size of $a$ in period $t$, if there exists $i$ such that $\min_i + \max_i >= a$ and $a > \min_i$, accept this arrival at row $i$, go to step 5; otherwise, go to step 4.
    \item[Step 4.] Find a maximal group of seats to accept this arrival, otherwise, reject this arrival.
    % Use the information of realized arrivals to make the decision.
    \item[Step 5.] Move to the next period. Repeat step 3. 
  \end{description}
\end{algorithm}

Step 2: the pattern will have the same loss by these procedures. ($\min_i$ can be 0.)
(Can we use the partial information?)

% \begin{lem}
%   Any largest patterns can be generated by the largest pattern constructed from the greedy method.
% \end{lem}

This method can be used without stochastic information. The performance will improve when the total demand can construct the largest patterns for all rows.


% Once: Obtain the supply from the stochastic model by benders decomposition. Use the deterministic model to obtain a heuristi supply. Then use the multi-class rule to decide whether to accept the group at each period.

% Several: Initially, set the mean demand for all periods as the upper bound of demand. Then obtain the supply from the deterministic model. Set the accepted demand as the lower bound of demand, the upper bound of demand will be the sum of accepted demand and mean demand for the remaining periods. Update the lower bound and upper bound when some supply runs out.

% One counterexample: [15,21,13,3] /[15,21,10,3]  reject 4

% Different types of movies will have different probabilities, consider the preference for policy when demand = supply.


% The numbers in the `performance compared to the optimal' represent M1, M2, M3, M4, M5, M6 respectively in order.

% The maximal number of people served can be obtained by \eqref{deter_upper} with a realized sequence of arrival.

% Some important information:
% The government will give the restriction: 入座人数50% /相连座位 4个
