% !TEX root = sum1.tex
\section{Literature Review}

The present study is closedly connected to the following research areas- seat assignment with social distancing, group seat reservation and dynamic seat assignment. The subsequent sections review literature pertaining to each perspective and highlight significant differences between the present study and previous research.

\subsection{Seat assignment with social distance}

While social distancing is a recognized and practiced method for containing the spread of infectious diseases \cite{moosa2020effectiveness}, operational guidance for its implementation is still lacking, particularly in social distance measures that involve operational details. An example of such a measure is ensuring social distancing in seating plans.

Social distancing in seat assignment has attracted considerable attention from the research coterie since pandemics.  \cite{bortolete2022support} presents a tool for classroom layout planning that considers social distancing during the COVID-19 pandemic. The tool addresses both fixed and non-fixed position seat allocation problems, utilizing integer optimization models and circle packing techniques.

\cite{ghorbani2020model} proposes a model to minimize the economic impact of COVID-19 on businesses that have implemented social distancing measures, while also minimizing the health impact on customers and employees. The study applies the model to the scenario of aircraft seat layouts using the Annealing Monte Carlo technique, demonstrating the necessity of using the model to find optimal arrangements of passengers.

% The COVID-19 pandemic has led to strict regulations that have affected major economies, prompting policymakers and planners to devise strategies that balance virus control with economic activities.

\cite{haque2022optimization} discusses the potential implications of seat inventory management in long-distance passenger trains and proposes a novel seat assignment policy that aims to reduce virus diffusion risk among passengers by minimizing interaction. A mixed-integer linear programming problem is formulated to simultaneously maximize operator revenue and minimize virus diffusion, and tested using real-life data from Indian Railways. 

\cite{fischetti2021safe} proposes a parallel between this problem and that of locating wind turbines in an offshore area, allowing for the application of Mathematical Optimization algorithms to produce optimized facility layouts that minimize the overall risk of infection among customers. 

\cite{haque2023social} proposes a novel seating assignment policy that minimizes virus spread risk while maximizing revenue for a train operating company. The proposed mixed-integer programming model considers several factors that govern virus spread and includes valid inequalities and preprocessing steps to strengthen the formulation. The model is evaluated using real-life data from Indian Railways, demonstrating a significant reduction in the risk of contagion and an increase in seat utilization compared to the current approach employed by operators.

However, these researchs focus on the static version of the problem. This typically involves creating an IP model with social distancing constraints, which is then solved either heuristically or directly. However, the social distancing constraint in their works- which requires individuals to maintain a certain distance from one another - means that a fixed seating layout cannot fully utilize available space and may not be practical for commercial use.

Meanwhile, for seat assignment during booking period, once the group is accepted, the seats should be assigned and will not be changed. 


As we transition into the post-pandemic era, many people may still be concerned about the risk of contracting the virus. This may lead to hesitancy in attending social events or gatherings that require close proximity to others.

Maintaining social distancing can be challenging in crowded public spaces, which can cause further anxiety for individuals. Meanwhile, social distancing measures can have a significant economic impact, especially on businesses that rely on large crowds or close contact between individuals.

However, if organizers can develop a scheme that balances the need for social distancing with economic considerations, it could help to alleviate people's concerns and protect their health. Such a scheme would go a long way in dispelling people's doubts and fears. Although social distance has reduced some seat income, people have less potential concerns and are more willing to buy tickets.



% Almost all existing work focuses on the static version of seat assignment with social distance. They commonly build a IP model with social distance constraint, then solve it heuristically or directly.
% The social distance constraint is that people should keep a certain distance. However, for a fixed seating layout, the setting cannot fully utilize seating space and cannot be applied in commerce.


% Dynamic seat assignment with social distance can be implemented manually by the venue staff, or through automated systems that use algorithms to optimize the seating arrangements based on various factors, such as the number of available seats, customer preferences, and the recommended distance between audience members.

\subsection{Group seat reservation}
The group seat reservation policy, which ensures members of the same family or group are seated together, has various applications in industries such as hotels \cite{li2013modeling}, working spaces\cite{fischetti2021safe}, public transport\cite{deplano2019offline}, sports arenas\cite{kwag2022optimal}, and large-scale events \cite{lewis2016creating}. This policy has significant impacts on passenger satisfaction and revenue, with the study \cite{yuen2002group} showing that passenger groups increase revenue by filling seats that would otherwise be empty. Traditional works \cite{clausen2010off, deplano2019offline}in transportation focus on maximizing capacity utilization or reducing total capacity needed for passenger rail, typically modeling these problems as knapsack or binpacking problems.


In contrast, there is a lack of research on group seat reservations for booking tickets for cinemas, where available seats are typically displayed for customers to choose from for low-demand movie tickets. For concerts with high demand, it is usually not possible to choose seats independently, and the organizer will inform seat information after confirming the order. However, the COVID-19 pandemic has shed new light on the potential benefits of group reservations, as they can improve revenue without increasing the risk of infection.

% However, for booking tickets for cinemas or concerts, there are almost no articles discussing group seat reservations. For movie tickets with low demand, the system will display a table of available seats for customers to choose from. For concerts with high demand, it is usually not possible to choose seats independently. The organizer will inform the seat information after confirming the order.
% The recent pandemic has offered new insights on group reservations as they exhibit improvement in revenue without an increase in the risk of getting infected. 


Fischetti et al. \cite{fischetti2021safe} proposed a seating assignment policy for groups of customers entering restaurants or amphitheaters, with the objective of minimizing transmission risk by minimizing interactions. Dundar and Karakose \cite{dundar2021seat} proposed a two-stage algorithm for classroom seat assignment during the pandemic, with the first phase maximizing total allocations and the second phase maximizing the minimum interpersonal distance between students. Salari et al. \cite{salari2022social} performed group seat assignment in airplanes during the pandemic and found that increasing passenger groups can yield greater social distance than single passengers. Similarly, Moore et al. \cite{moore2021seat} suggest that relaxing social distancing norms within a family may increase occupancy without an significant increasing in virus transmission risk.


% For the sake of convenience, we list some of the intricate characteristics of the related studies and our study in Table 1.


\subsection{Dynamic seat assignment}
Dynamic seat assignment is a process of assigning seats to passengers on a transportation vehicle, such as an airplane, train, or bus, in a way that maximizes the efficiency and convenience of the seating arrangements \cite{hamdouch2011schedule, berge1993demand}. 

\cite{jiang2015dynamic} proposes a revenue management approach for high-speed rail (HSR) passenger ticket assignment with dynamic adjustments. The approach integrates short-term demand forecasting, ticket assignment, and dynamic ticket adjustment mechanisms to allocate passenger tickets during presale periods and avoid situations where tickets are insufficient at some stations while seats remain empty.

Dynamic capacity management in airlines involves reassigning aircraft to match the supply and demand for seats, resulting in increased passenger capacity and revenue. \cite{de2004impact} proposes an EMSRd derivative of the EMSRb booking limit calculation method that considers the impact of future capacity changes. The EMSRd approach efficiently solves the fleet assignment problem and shows potential revenue improvements of over 1\%.

\cite{zhu2023assign} addresses a revenue management problem arising from the selling of high-speed train tickets in China, where each request must be assigned to a single seat throughout the entire journey. The authors propose a modified network revenue management model and introduce a bid-price control policy based on a novel maximal sequence principle. They also propose a "re-solving a dynamic primal" policy that achieves uniformly bounded revenue loss. The study reveals connections between this problem and traditional network revenue management problems and shows that the impact of the assign-to-seat restriction can be limited with the proposed methods.

Dynamic seat assignment is also used in cinemas and concert venues to optimize the seating arrangements for the audience, with the goal of providing the best possible experience for each audience member while maximizing ticket sales and revenue for the venue.

However, implementing dynamic seat assignment with social distance in cinemas and concert venues presents several challenges. The primary challenge is to balance the need for safety with the desire to maximize revenue for the venue. This requires careful consideration of various factors, such as the layout of the venue, the number of available seats, and the preferences of the audience. To implement dynamic seat assignment with social distance, venues may need to adjust their seating plans to accommodate the recommended distance between audience members. This may involve reducing the overall number of seats or reconfiguring the seating arrangements.

Implementing dynamic seat assignment with social distance can be done manually by the staff or through automated systems that use algorithms to optimize the seat assignments based on various factors, such as ticket sales, seat availability, and customer preferences. However, the implementation of social distance measures poses unique challenges that require careful planning and consideration of various factors to balance safety with revenue generation.

Despite these challenges, dynamic seat assignment with social distance is a promising approach to help ensure the safety of audience members in cinemas, concert venues, and other public spaces during the COVID-19 pandemic.

Dynamic Knapsack Problem? Revenue management?



% \subsection{Scenario generation}

% It is challenging to consider all the possible realizations; thus, it is practicable to use discrete distributions with a finite number of scenarios to approximate the random demands. This procedure is often called scenario generation.

% Some papers consider obtaining a set of scenarios that realistically represents the distributions of the random parameters but is not too large. \cite{feng2013scenario} \cite{casey2005scenario}
% \cite{henrion2018problem}

% Another process to reduce the calculation is called scenario reduction. It tries to approximate the original scenario set with a smaller subset that retains essential features.



% Every time we can regenerate the scenario based on the realized demands. (Use the conditional distribution or the truncated distribution)


% Suppose that the groups arrive from small to large according to their size. Once a larger group comes, the smaller one will never appear again.

% When a new group arrives (suppose we have accepted $n$ groups with the same size), we accept or reject it according to the supply (when $n+1 < \text{supply}$, we accept it). 
% then update the scenario set according to the truncated distribution. We can obtain a new supply with the new probability and scenario set.


\newpage
